---
title: "F Ratio - OLS regression"
author: "Sri Seshadri"
date: "1/20/2018"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question

Page 76 of ISLR states that "...When there is no relationship between the response and predictors, one would expect the F-Statistic to take on a value close to 1. On the other hand, If $H_a$ is true then E{(TSS - RSS)/p} > $\sigma^2$ , so we expect F to be greater than 1 "

**If there was no relationship between the predictor and the response variable shouldn't the F statistic be very less than on equal to 1? instead of close to 1?** 

# Experiment

Generate random numbers and randomly assign the levels of the predictor variable to the random numbers. Fit a OLS regression and inspect the ANOVA table for F ratio. Repeat the experiemt for 1000 iterations.

```{r}
getFStats <- function() {
  # Generate random numbers from a standard normal distribution
  
  y <- rnorm(1000)
  
  # generate predictor variable with levels 1 through 10 
  
  x <- rep(1:10,100)
  
  # plot y vv x
  
  #plot(y~x)
  
  # Simple linear regression
  
  model.lm <- lm(formula = y ~ x)
  
  #summary(model.lm)
  
  Fstats <- anova(model.lm)$`F value`[1]
  return(Fstats)
}

result <- replicate(1000,getFStats())

hist(result,breaks = 20,col = "blue",xlab = "F statistic",main = "Histogram of F Ratios")
abline(v=1,col = "red",lty = 5)
```

```{r}
f<-seq(.1,20,.1)
plot(ecdf(result),col='blue',xlab='F-statistic',ylab='CDF',main='F-statistic CDFs',lty=1)
lines(f,pf(f,1,998),col="red",type='l',lty=2)
legend('center',legend=c('True F(1,998) CDF','Empirical F-statistics CDF'),col=c('Red','Blue'),lty=c(2,1),box.col=NA)
abline(v=1,lty=3)
```


# Conclusion

The quote in ISLR is misleading and should read "F statistic should take on values less than on equal to 1".
I am open to criticism, please point me in a direction should you feel the experimentation is not set up correctly. 

