---
title: "Lab6"
author: "Sri Seshadri"
date: "2/17/2018"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(ISLR)
library(dplyr)
library(broom)
```

## Lab 6 - Logistic regression, LDA, QDA and KNN


##  4.6.1 The stock Market Data

Dataset used Smarket from the ISLR library; the data contains

  - For each date, % returns of each of the five previous days were recorded.(Lag 1 thru 5)
  - Volume - the number of shares traded on the previous day in billions
  - Today - the % return on the date in question
  - Direction - market went up or down on this date?

## Data inspection {.smaller .flexbox .vcenter}

```{r cars, echo = TRUE,collapse=TRUE}
Smarket <- Smarket
skimr::skim(Smarket)
```

## Explore variables

```{r Smarket}
pairs(Smarket,cex = 0.5, col = c("grey"))
```

## Correlations

```{r, echo=T}
numericcols <- sapply(Smarket,is.numeric)
correlations <- cor(Smarket[,numericcols])
corrplot::corrplot(correlations,method = "ellipse")
```

## Correlations with Direction {.smaller}

```{r, echo=T}
Smarket$Direction <- as.numeric(Smarket$Direction)
par(mfrow = c(1,2))
corrplot::corrplot(cor(Smarket), method = 'ellipse')
beanplot::beanplot(Volume~Year, data = Smarket,lab = "year", ylab = "Volume", col = "grey")
```

## 4.6.2 Logistic Regression {.smaller}

*Remember*  The model fits the log odds or logit. The output need do be "unlogitted"
```{r, echo=T}
# Make Direction a factor variable again; simply reload the data from package
Smarket <- ISLR::Smarket
glm.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume , family = 'binomial',data = Smarket)
summary(glm.fit)

# significance of fit
1- pchisq(glm.fit$null.deviance - glm.fit$deviance,glm.fit$df.null - glm.fit$df.residual)

```

## Predict() usage

```{r, echo=T}
glm.probs <- predict(glm.fit, type = "response")
head(glm.probs)

# probability of direction going up or down? What did R do... it created a dummy variable with Up as 1 and down as zero
contrasts(Smarket$Direction)
```

## Confusion Matrix {.smaller}

```{r}
glm.pred <- rep("Down", nrow(Smarket))
glm.pred[glm.probs > 0.5] <- "Up"
# now the confusion matrix
caret::confusionMatrix(glm.pred,reference = Smarket$Direction, positive = "Up")
```

## Leave 2005 data as hold out set {.smaller}

```{r}
train <- Smarket$Year < 2005
glm.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Smarket, family = 'binomial',subset = train)
summary(glm.fit)

glm.probs <- predict(glm.fit, newdata = Smarket[!train,], type = "response")

glm.pred <- rep("Down", nrow(Smarket[!train,]))
glm.pred[glm.probs > 0.5] <- "Up"

caret::confusionMatrix(glm.pred, reference = Smarket$Direction[!train], positive = "Up")
```

## Use Lag 1 & 2 as predictors {.smaller}

Since Lag 1 & Lag 2 had the least p values, lets try using those alone as predictors.

```{r}
glm.fit <- glm(Direction ~ Lag1 + Lag2, data = Smarket, family = 'binomial',subset = train)
summary(glm.fit)

glm.probs <- predict(glm.fit, newdata = Smarket[!train,], type = "response")

glm.pred <- rep("Down", nrow(Smarket[!train,]))
glm.pred[glm.probs > 0.5] <- "Up"

caret::confusionMatrix(glm.pred, reference = Smarket$Direction[!train], positive = "Up")
```

## Prediction for certain lags

```{r}
predict(glm.fit,newdata = data.frame(Lag1 = c(1.2,1.5), Lag2 = c(1.1,-0.8)), type = "response")
```

## 4.6.3 Linear Discriminant Analysis {.smaller}

LDA is best when the assumption of gaussian distribution of X is satisfied and equality of covariance across predictors.

```{r}
library(MASS)
lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket,subset = train)
lda.fit
```

## Plot of LDA model {.smaller}
What's been plotted? 
coef of Lag1 * Lag 1 + coef of Lag2 * lag2
What does this mean?

```{r,echo=T}
plot(lda.fit)
```

## Reproduce plot(lda.fit) {.smaller}


```{r, echo=T}
test <- coef(lda.fit)[1] %*% Smarket[train,"Lag1"] + 
        coef(lda.fit)[2] %*% Smarket[train,"Lag2"]
Direc <- Smarket[train,"Direction"]
par(mfrow = c(1,2))
hist(test[Direc=="Up"],freq = F)
hist(test[Direc=="Down"], freq = F)
```

